Selecting Computations: Theory and Applications

### 2012-03-30 paper/vago.uai/context.tex
added files, probably too many

\note{No index theorem; via context inversion counter-example}

de

\begin{figure}[htb]
\centering
\includegraphics[scale=0.7, trim=90 70 400 300]{swap-counterex.pdf}
\caption{Regret vs. cost}
\label{fig:swap-counterex}
\end{figure}

@@

set terminal postscript enhanced linewidth 2
set size 0.5,0.5
set output "swap-counterex.ps"
set xlabel "Utility of fixed alternative (u)"
set ylabel "Utility relative to stopping"
set xtics 0.5
set ytics 0.1
set key left top
plot [u=-2:2] [-0.3:0.3] 0 title "stop" with lines lw 2, qa(u)-q0(u) title "observe A" with lines lw 2, qb(u)-q0(u) title "observe B" with lines lw 2	
\end{comment}

@@

@@

\begin{thm}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, 
	there exists a real interval $I(s)$ for every state $s\in I$ such that
	it is optimal to stop in state $s$ in $M_\mu$ iff $\mu\in I(s)$.
	Furthermore, $I(s)$ contains $\max_i\mu_i(s)$ whenever it is nonempty.
\end{thm}

\begin{proof}

With $m(s) = \max_i\mu_i(s)$, the utility of a policy $\pi$ starting in state $s$ of $M_\mu$ is
\[
	V^\pi_{M_\mu}(s) = \IE_{\pi}[-c\,N + \max(\mu,m(S_N)) \given S_0=s]
\]
and the utility of stopping in this state $\max(\mu,m(s_0))$.
We wish to show that the set of $\mu$ such that
\[
	\max_\pi \IE_{\pi}[-c\,N + \max(\mu,m(S_N)) - \max(\mu,m(S_0)) \given S_0=s] \le 0
\]
forms an interval.

@@

Applying this with $v_1 = m(s_0)$ and $v_2=\IE_{\pi}[c\,N]$, and observing that the union of intervals containing
a point is an interval containing that point, gives the result.
\end{proof}

>>>>

%% \note{No index theorem; via context inversion counter-example}

The Gittins index theorem \citep{Gittins:1979} is a famous structural
result for bandit problems.  It states that in bandit problems with
independent reward distribution for each arm and geometric discounting,
the optimal policy is an \term{index policy}:  each arm is assigned a
real-valued index based on its state only, such that it is optimal to
sample the arm with greatest index.

The analogous result does \emph{not} hold for metalevel decision problems,
even when the action's values are independent (this formalized later in \dfnref{dfn:independent-actions}):

\begin{example}[Non-indexability]
Consider a metalevel probability model with three actions. 
	$U_0=m$ has a known value (the context),
	$U_1$ is equally likely to be -1.5 or 1.5 (low mean, high variance),
	and $U_2$ is equally likely to be 0.25 or 1.75 (high mean, low variance).
The two computations are to observe exactly $U_1$ and $U_2$, respectively, with a cost 0.2 of computation.
The corresponding metalevel MDP has 9 states and can be solved exactly.
\figref{fig:swap-counterex} plots $Q^*_m(s_0,U_i) - Q^*_m(s_0,\bot)$ as a function of the known value $m$.
As the context $m$ varies the optimal action inverts from observing 1 to observing 2.
Inversions like this are impossible for index policies.
\end{example}

\begin{figure}[htb]
\centering
\includegraphics[scale=0.7, trim=90 70 400 300]{swap-counterex.pdf}
\caption{Optimal Q-values as a function of the utility of the fixed alternative.  Note the inversion
where for low $m$ observing action~$1$ is strictly optimal, while for medium $u$ observing action~$2$ is strictly optimal.}
\label{fig:swap-counterex}
\end{figure}

@@

set terminal postscript enhanced linewidth 2
set size 0.5,0.5
set output "swap-counterex.ps"
set xlabel "Utility of fixed alternative (m)"
set ylabel "Utility relative to stopping"
set xtics 0.5
set ytics 0.1
set key left top
plot [u=-2:2] [-0.3:0.3] 0 title "stop" with lines lw 2, qa(u)-q0(u) title "observe 1" with lines lw 2, qb(u)-q0(u) title "observe 2" with lines lw 2	
\end{comment}

@@

There is, however, a restriction on what kind of influence the context can have:

@@

\begin{thm}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, 
	there exists a real interval $I(s)$ for every state $s\in I$ such that
	it is optimal to stop in state $s$ in $M_\mu$ iff $\mu\notin I(s)$.
	Furthermore, $I(s)$ contains $\max_i\mu_i(s)$ whenever it is nonempty.
\end{thm}

\begin{comment}
	\begin{proof}
	With $m(s) = \max_i\mu_i(s)$, the utility of a policy $\pi$ starting in state $s$ of $M_\mu$ is
	\[
		V^\pi_{M_\mu}(s) = \IE_{\pi}[-c\,N + \max(\mu,m(S_N)) \given S_0=s]
	\]
	and the utility of stopping in this state $\max(\mu,m(s_0))$.
	We wish to show that the set of $\mu$ such that
	\[
		\max_\pi \IE_{\pi}[-c\,N + \max(\mu,m(S_N)) - \max(\mu,m(S_0)) \given S_0=s] \le 0
	\]
	forms an interval.

@@

	Applying this with $v_1 = m(s_0)$ and $v_2=\IE_{\pi}[c\,N]$, and observing that the union of intervals containing
	a point is an interval containing that point, gives the result.
	\end{proof}	
\end{comment}

### 2012-03-30 paper/vago.uai/dropped-parts.tex
added files, probably too many

%% \section{On optimal policies for selection}

%% Define selection problem and metalevel MDP representation.

%% \note{Unsure about terminology here: selection problems, sampling problems, 
%% metalevel probability model, metalevel decision problem (conflicts with Markov decision process?).}

%% ============================
\subsection{Selection problems}
%% ============================

\note{Formal definition of selection problems and the metalevel MDP with cost per sample (time value); also mention budgeted learning.}

In a \term{selection problem} the decision maker is faced with the choice between
$k$ alternative actions.  To make this choice, they may gather evidence about the
utility of each of these alternative, at some cost.  What the decision maker can know
and what they can find out is formalized in:

\begin{dfn} \label{dfn:metalevel-model}
	A \term{metalevel probability model} is a tuple $M=(U_1,\dots,U_k,\Evidence)$ 
	consisting of jointly distributed random variables:
	\begin{itemize}
		\item Real random variables $U_1,\dots,U_k$, where $U_i$ is the utility of performing action~$i$, and
		\item A countable set $\Evidence$ of random variables, each variable $E\in\Evidence$ being 
		      a computation that can be performed and whose value is the result of that computation.
	\end{itemize}
\end{dfn}
%% Not actually true that we necessarily define this with U's as roots:  in sampling models they are naturally /leaves/.

A metalevel probability model, when combined with a cost $c>0$ of computation,%
\footnote{The assumption of a fixed cost of computation is a simplification; 
	precise conditions for its validity are given by \citep{Harada:1997}.} 
defines a metalevel decision problem: what is the optimal strategy with which to choose a sequence 
of computations $E\in\Evidence$ to observe in order to maximize the expected utility 
of the action finally taken less the cost of computation?  Intuitively, 
this strategy should observe the variables which give the most evidence relevant
to deciding which action to perform, stopping when the cost of computation 
outweighs the benefit gained.

\begin{example}[Bernoulli sampling]\label{example:bernoulli}
One simple metalevel probability model is used as a model of the results of simulations \cite{someone}.
Each action will either succeed or not $U_i\in\{0,1\}$, with an unknown latent frequency of success $\Theta_i$, 
and stochastic simulations of possible consequences $E_{ij}$ can be performed:
\begin{align*}
	\Theta_i &\iid {\rm Uniform}[0,1]                      &&\quad\text{for $i=1,\dots,k$}\\
	U_i \given \Theta_i &\sample {\rm Bernoulli}(\Theta_i) &&\quad\text{for $i=1,\dots,k$}\\
	E_{ij} \given \Theta_i &\iid {\rm Bernoulli}(\Theta_i) &&\quad\text{for $i=1,\dots,k$ and $j=1,\dots$}
\end{align*}
\end{example}

%% Bounedness assumption
For simplicity, in the below we'll assume the utilities $U_i$ are bounded and so, 
without loss of generality, lie in $[0,1]$.

To formalize this problem, we use the theory of Markov Decision Processes (MDPs; 
see, for example, \citet{Puterman:1994}):

\begin{dfn}
A (countable state, undiscounted) \term{Markov Decision Process} (MDP) is a tuple $M=(S,s_0,A_s,T,R)$ where:
	$S$ is a countable set of states,
	$s_0\in S$ is the fixed initial state,
	$A_s$ is a countable set of actions available in state $s\in S$,
	$T(s,a,s')$ is the transition distribution 
	giving the probability of transitioning from a state $s\in S$ to a state $s'\in S$ after performing action $a\in A_s$,
	and $R(s,a,s')$ is the expected reward received when transitioning from $s\in S$ to $s'\in S$ after performing action $a\in A_s$.
\end{dfn}

\begin{dfn}\label{dfn:metalevel-mdp}
	Given a metalevel probability model%
		\footnote{Definition~\ref{dfn:metalevel-model} made no assumption about the computational result
			variables $E_i\in\Evidence$, but for simplicity in the following we'll assume that
			each $E_i$ takes one of a countable set of values.  Without loss of generality, 
			we'll further assume the domains of the computational variables $E\in\Evidence$ are disjoint.}
	$M=(U_1,\dots,U_k,\Evidence)$ and
	a cost of computation $c>0$, a corresponding \term{metalevel decision problem}
	is the MDP $M=(S,s_0,A_s,T,R)$ with
	\begin{align*}
		S &= \{\bot\}\cup\{\langle e_1\dots, e_n \rangle : e_i\in E_i \text{ for all $i$,} \\
								& \qquad\qquad \text{for finite $n\ge0$ and distinct $E_i\in\Evidence$}\} \\
		s_0 &= \langle \rangle \\
		A_s &= \{\bot\}\cup\Evidence_s \\
	%
	\intertext{where $\bot\in S$ is the unique terminal state,
	where $\Evidence_s\subseteq\Evidence$ is a potentially state-dependent subset of allowed computations,
	and when given any $s=\langle e_1, \dots, e_n \rangle\in S$,
	computational action $E\in\Evidence$, 
	and $s'= \langle e_1, \dots, e_n, e \rangle\in S$ where $e\in E$, we have:}
	%
		T(s,E,s') &= P(E=e \given E_1=e_1,\dots,E_n=e_n) \\
		T(s,\bot,\bot) &= 1 \\
		R(s,E,s') &= -c \\
		R(s,\bot,\bot) &= \max_i \mu_i(s) % \max_i \IE[U_i \given E_1=e_1,\dots,E_n=e_n]
	\end{align*}
	where $\mu_i(s) = \IE[U_i \given E_1=e_1,\dots,E_n=e_n]$.
\end{dfn}

\begin{example}[Bernoulli sampling]\label{example:bernoulli2}
In the Bernoulli metalevel probability model (\exampleref{example:bernoulli}),
note that: 
\begin{align}
	\Theta_i \given E_{i1},\dots,E_{in_i} &\sim {\rm Beta}(s_i+1, f_i+1)  \label{eq:bernoulli1}\\
	E_{i(n_i+1)} \given E_{i1},\dots,E_{in_i} &\sim {\rm Bernoulli}\left(\frac{s_i+1}{n_i+2}\right) \label{eq:bernoulli2} \\
	\IE[U_i \given E_{i1},\dots,E_{in_i}] &= (s_i+1)/(n_i+2) \label{eq:bernoulli3}
\end{align}
by standard properties of these distributions, where $s_i=\sum_{j=1}^{n_i} E_{in_i}$
is the number of simulated successes of action $i$, and $f_i=n_i-s_i$ the failures.  By \eqref{eq:bernoulli1}, 
the state space is the set of all $k$ pairs $(s_i,f_i)$, and \eqrefs{eq:bernoulli2}{eq:bernoulli3}
suffice to give the transition probabilities and terminal rewards, respectively.
\end{example}

Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ one defines policies and 
value functions as in all MDPs.  A (deterministic, stationary) \term{metalevel policy} 
$\pi$ is a function mapping states $s\in S$ to actions to take in that state $\pi(s)\in A_s$.
The \term{value function} for a policy $\pi$ gives the expected total reward
received under that policy starting from a given state $s\in S$: 
\begin{equation}\label{eq:value-fn}
	V^\pi_M(s) = \IE^\pi_M\left[ \sum_{i=0}^N R(S_i,\pi(S_i),S_{i+1}) \given S_0 = s\right]	
\end{equation}

where $N\in[0,\infty]$ is the random time the MDP is terminated, i.e.,
the unique time where $\pi(S_N)=\bot$.%
\footnote{Whenever $N<\infty$, $S_{N+1}=\bot$ the unique terminal state.
Instead of having a random termination time, one can equivalently make
the state $\bot$ absorbing with zero reward on all transitions.}

\begin{thm}\label{thm:value-of-computation}
	The value function of a metalevel decision process $M=(S,s_0,A_s,T,R)$ is of the form
	\[
		V^\pi_M(s) = \IE^\pi_M[ -c\,N + \max_i \mu_i(S_N) \given S_0=s]
	\]
	where the random variable $N$ denotes the total number of computations performed.
\end{thm}

\begin{proof}
	Follows immediately from \eqref{eq:value-fn} and the definition of the
	reward function $R$ in \dfnref{dfn:metalevel-mdp}.
\end{proof}

\begin{thm}
	MDPs $M=(S,s_0,A_s,T,R)$ where there exists a unique terminal state $\bot\in S$,
	a distinguished action $\bot\in A_s$ for all $s\in S$, a constant $c>0$, and bounded functions 
	$\mu_1,\dots,\mu_k\maps S\to\R$ such that for all $s\in S$, $a\in A_s\setminus\{\bot\}$, and $i=1,\dots,k$:
	\begin{align*}
		\mu_i(s) &= \sum_{s'\in S} T(s,a,s')\,\mu_i(s') \\
		T(s,\bot,\bot) &= 1 \\
		R(s,a,s') &= -c \\
		R(s,\bot,\bot) &= \max_i \mu_i(s)
	\end{align*}
	are equivalent to metalevel decision problems with bounded utilities, and conversely any 
	metalevel decision problem with bounded utilities $U_i$ is such an MDP.
\end{thm}

@@

For the converse, observe that the properties hold with $c$ the cost of computation
and \[\mu_i(s) = \IE(U_i\given E_1=e_1,\dots,E_n=e_n)\] where $s=\langle e_1,\dots,e_n \rangle$.
\end{proof}

As usual, an \term{optimal policy} $\pi^*$ is one which maximises 
the value from every state $s\in S$, i.e., if we define for each $s\in S$
\[
	V^*_M(s) = \sup_\pi V^\pi_M(s),
\]
then an optimal policy $\pi^*$ satisfies $V^{\pi^*}_M(s) = V^*_M(s)$
for all $s\in S$.

\note{Something about existence of optimal solutions?  Probably do not exist in general,
but do for finite $S$ despite discounting as this is a positive bounded model.}

%% Positive bounded model.  We do have upper and lower bounds on V^*.  For finite s we can get optimality...

\begin{dfn}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
	the \term{myopic policy} $\pi^m(s)$ is defined to equal
	\begin{equation}\label{eq:myopic}
		\argmax_{E\in A_s} \IE_M[ -c + \max_i \mu_i(S_1) \given S_0 = s, A_0 = E]		
	\end{equation}
	whenever this maximum is positive, $\pi^m(s)=\bot$ otherwise.
\end{dfn}

The myopic policy takes the best action, to either stop or perform a computation,
under the assumption that at most one further computation can be performed.  This policy
is typically easy to compute, although not necessarily very good.  It does have
importance, however, through its relation to the optimal policy in \thmref{thm:optimal-myopic}
and its partial converse \thmref{thm:myopic-optimal}:

%% \note{Theorem: if Optimal stops in x, myopic stops in x (converse is more useful!)}

\begin{thm}\label{thm:optimal-myopic}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$
	if the optimal policy stops in a state $s\in S$
	then the myopic policy stops too, i.e., if $\pi^*(s)=\bot$ then $\pi^m(s)=\bot$.
\end{thm}

\begin{proof}
	If the optimal policy stops in a state $s\in S$ then
	\[
		V^{\pi^*}(s) \le \max_i \mu_i(s),
	\]
	but then the maximum of \eqref{eq:myopic} is non-positive, so the 
	myopic policy stops by definition.
\end{proof}

@@

\begin{proof}
Defining $m(s) = \max_i\mu_i(s)$, observe the myopic stopping for all
reachable states implies that
\begin{align*}
	\IE^{\pi}[m(S_{j+1}) - c; j<N\given S_0=s] \\
	\le \IE^{\pi}[m(S_{j}); j<N\given S_0=s]
\end{align*}
holds for all $j$, and as a result:
\begin{align*}
	V^\pi(s) 
	&= \IE^{\pi}[ - c N + m(S_N) |S_0=s] \\
	&= \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} (m(S_{j+1}) - c - m(S_j)) |S_0=s] \\	
	&\le \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} 0 |S_0=s] \\		
	&= \max_i\mu_i(s)
\end{align*}
\end{proof}

% Similarity to an optimal stopping result
	% Search <blah> for one-step lookahead rules.
	% But HERE is generalized to a stopping problem where there are many ways to continue.

\note{Define value of perfect information here for the below?}

\begin{thm}\label{thm:bounded-expected-computations}
	The optimal policy $\pi^*$ has bounded expected number of computations
	\[
		\IE^{\pi^*}[N\given S_0=s] \le \frac{1}{c} \left(\IE[\max_i U_i\given S_0=s] - \max_i \mu_i(s)\right).
	\]
	Further, any policy $\pi$ with infinite expected number of computations 
	$\IE^\pi[N]=\infty$ has negative infinite value, so in particular the optimal
	policy stops with probability 1.
\end{thm}

\begin{proof}
	The first follows as in state $s$ the optimal policy has value at least that
	of stopping immediately $\max_i \mu_i(s)$, and as $\IE \max_i\mu_i(S_N) \le \IE \max_i U_i$ by Jensen's inequality.
	The second follows from \thmref{thm:value-of-computation} as $c>0$.
\end{proof}

Although the \emph{expected} number of computations is always bounded,
there are important cases in which the \emph{actual} number is not:

\begin{example}\label{example:sprt}
Consider the Bernoulli sampling model for two actions but with different priors:
	$\Theta_1=1/2$,
	and $\Theta_2$ is $1/3$ or $2/3$ with equal probability.
%
Simulating action 1 gains nothing, and after $(s,f)$ simulated successes and failures of action 2
the posterior odds is
\[
	\frac{P(\Theta_2=2/3\given s,f)}{P(\Theta_2=1/3\given s,f)} = 2^{s-f}.
\]
The difference $s-f$ is a sufficient statistic for $\Theta$ given the computations,
so there exists a metalevel MDP with this difference as its state.  For sufficiently
low cost, the optimal policy samples in states $-1$, $0$, and $1$.  But with probability
$1/3$ state $0$ transitions to state $0$ after two samples, giving finite probability
to arbitrarily long sequences of computations.
\end{example}

However in a number of settings, including the original Bernoulli model of \exampleref{example:bernoulli},
we can upper bound the number of computations:

\begin{thm}\label{thm:bernoulli-bound}
In the Bernoulli sampling problem of \exampleref{example:bernoulli},
at most $k/4c$ computations are performed.	
\end{thm}

\begin{proof}

\end{proof}

As an important consequence, we can truncate the MDP to

\note{Gaussian bound too?}

%% \note{Reference ranking and selection here?}

The bound in \thmref{thm:bernoulli-bound} allows us

\begin{dfn}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
	the \term{blinkered policy} $\pi^b(s)$ \term{myopic policy} $\pi^m(s)$ is defined to equal
	\begin{equation}\label{eq:blinkered}
		V^b(s) = \sup_{\pi\in\Pi_b} V^\pi_M(s),
	\end{equation}
	where $\Pi_b$ is the set of policies which sample .....

	...need comparisons between $A_s$'s to exist.
\end{dfn}

\note{Graph comparing blinkered to other policies}

\begin{figure}[htb]
\centering
\includegraphics[scale=0.7, trim=90 70 400 300]{blinkered-regret.pdf}
\caption{Regret vs. cost}
\label{fig:blinkered}
\end{figure}

\begin{comment}
set xlabel "Cost"
set ylabel "Regret"

set terminal postscript enhanced linewidth 2
set output "blinkered-regret.ps"
set size 0.5, 0.5

set logscale xy

plot [0.1:0.0005] [0.05:1] "blinkered-regret.dat" title "Blinkered" with lines lw 2, "espb-regret.dat" title "ESPb" with lines lw 2, "ucb-blinkered-regret.dat" title "UCB1-B" with lines lw 2, "ucb-blinkered-table-regret.dat" with lines lw 2 title "UCB1-b"	
\end{comment}

>>>>

\begin{comment}
	\begin{lmm}\label{lmm:equivalent-mdp}
		Given MDPs $M^1=(S^1,s^1_0,A^1_s,T^1,R^1)$ and $M^2=(S^2,s^2_0,A^2_s,T^2,R^2)$,
		if there exists $f\maps S^1\to S^2$ such that for all $s,s'\in S^1$ and $a\in A_s$,
		we have $A^2_{f(s)} = A^1_s$ and
		\begin{align*}
			T^2(f(s),a,f(s')) &= T(s_2,E,s_2') \\
			\max_i \mu_i(s_1) &= \max_i \mu_i(s_2),\\
		%
		\intertext{then for any policy $\pi^2$ in $M^2$, with policy $\pi^1$ in $M^1$
		defined by $\pi^1(s) = \pi^2(f(s))$ we have}
		%
			V^{\pi^1}_{M^1}(s) &= V^{\pi^2}_{M^2}(f(s)).
		\end{align*}
	\end{lmm}		
\end{comment}

\begin{comment}
	\begin{thm}
		MDPs $M=(S,s_0,A_s,T,R)$ where there exists a unique terminal state $\bot\in S$,
		a distinguished action $\bot\in A_s$ for all $s\in S$, a constant $c>0$, and bounded functions 
		$\mu_1,\dots,\mu_k\maps S\to\R$ such that for all $s\in S$, $a\in A_s\setminus\{\bot\}$, and $i=1,\dots,k$:
		\begin{align*}
			\mu_i(s) &= \sum_{s'\in S} T(s,a,s')\,\mu_i(s') \\
			T(s,\bot,\bot) &= 1 \\
			R(s,a,s') &= -c \\
			R(s,\bot,\bot) &= \max_i \mu_i(s)
		\end{align*}
		are equivalent to metalevel decision problems with bounded utilities, and conversely any 
		metalevel decision problem with bounded utilities $U_i$ is such an MDP.
	\end{thm}

@@

	For the converse, observe that the properties hold with $c$ the cost of computation
	and \[\mu_i(s) = \IE(U_i\given E_1=e_1,\dots,E_n=e_n)\] where $s=\langle e_1,\dots,e_n \rangle$.
	\end{proof}
\end{comment}

%% \note{Something about existence of optimal solutions?  Probably do not exist in general,
%% but do for finite $S$ despite discounting as this is a positive bounded model.}
%% Positive bounded model.  We do have upper and lower bounds on V^*.  For finite s we can get optimality...

%% \note{Define value of perfect information here for the below?}

@@

\begin{proof}
Defining $m(s) = \max_i\mu_i(s)$, observe the myopic stopping for all
reachable states implies that
\begin{align*}
	\IE^{\pi}[(m(S_{j+1}) - c)\, 1(j<N)\given S_0=s] \\
	\le \IE^{\pi}[m(S_{j})\, 1(j<N)\given S_0=s]
\end{align*}
holds for all $j$, and as a result:
\begin{align*}
	V^\pi(s) 
	&= \IE^{\pi}[ - c N + m(S_N) |S_0=s] \\
	&= \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} (m(S_{j+1}) - c - m(S_j)) |S_0=s] \\	
%%	&\le \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} 0 |S_0=s] \\		
	&\le \max_i\mu_i(s) \qedhere
\end{align*}
\end{proof}

\begin{thm}\label{thm:closed}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
	a subset $S'\subseteq S$ of states closed under transitions,
	and a computation $E\in\Evidence$,
	if $Q^m(s,E) \le Q^m(s,\bot)$ for all states $s\in S'$ where $E\in A_s$,
	then $Q^*(s,E) \le \max{E\neq a\in A_s} Q^*(s,a)$ for all states $s\in S'$ where $E\in A_s$.
\end{thm}

The state space in definition \dfnref{metalevel-mdp} is often a bit cumbersome for analysis
of particular examples, with many redundant.  However, we can use the following to reduce 
the MDP to something more manageable:
\begin{lmm}\label{lmm:equivalent-mdp}
	Given MDPs $M^1=(S^1,s^1_0,A^1_s,T^1,R^1)$ and $M^2=(S^2,s^2_0,A^2_s,T^2,R^2)$,
	if there exists $f\maps S^1\to S^2$ such that for all $s,s'\in S^1$ and $a\in A_s$,
	we have $A^2_{f(s)} = A^1_s$ and
	\begin{align*}
		T^2(f(s),a,f(s')) &= T(s_2,E,s_2') \\
		\max_i \mu_i(s_1) &= \max_i \mu_i(s_2),\\
	%
	\intertext{then for any policy $\pi^2$ in $M^2$, with policy $\pi^1$ in $M^1$
	defined by $\pi^1(s) = \pi^2(f(s))$ we have}
	%
		V^{\pi^1}_{M^1}(s) &= V^{\pi^2}_{M^2}(f(s)).
	\end{align*}
\end{lmm}

### 2012-03-30 paper/vago.uai/introduction.tex
added files, probably too many

The broad family of sequential decision problems includes
combinatorial search problems, game playing, robotic path planning,
model-predictive control problems, Markov decision processes (fully or
partially observable), and a huge range of applications. In almost all
realistic instances, exact solution is intractable and approximate
methods are sought.  Perhaps the most popular approach is to simulate
a limited number of possible future action sequences. For example,
a typical game-playing algorithm explores a tree or graph of action sequences
with some limit on depth, using pruning methods to avoid irrelevant
subtrees; based on what it finds in the explored portion of the state space, the algorithm
then selects a move.

Clearly, it is desirable to select the best possible move from the
least possible amount of exploration. For a given amount of
exploration, decision quality can be improved by directing exploration
towards those actions sequences whose outcomes are helpful in selecting
a good move. Thus, the {\em metalevel} decision problem is to choose
what future action sequences to explore (or, more generally, what
deliberative computations to do), while the {\em object-level}
decision problem is to choose an action to execute in the real world.

@@

An independent thread of research on metalevel control began with work
by \citet{Kocsis+Szepesvari:2006} on the UCT algorithm, which operates
in the context of {\em Monte Carlo tree search} (MCTS) algorithms.
In MCTS, each computation takes the form
of a simulating a randomized sequence of actions leading from a leaf of the
current tree to a terminal state. UCT is primarily a method for
selecting a leaf from which to conduct the next simulation.
It forms the core of the successful \textsc{MoGo} algorithm for Go playing \citep{Gelly+Silver:2011}.  It is situated 
within the theory of bandit problems \citep{Berry+Fristedt:1985}, applying the bandit algorithm
UCB1 \citep{Auer+et+al:2002} recursively to select actions to perform within simulations.

@@

Analyzing the metalevel problem within an appropriate theoretical
framework ought to lead to more effective algorithms than those
obtained within the bandit framework.  For Monte Carlo computations,
in which samples are gathered to estimate the utilities of actions,
the metalevel decision problem is an instance of the {\em selection
problem} studied in statistics~\cite{early-selection-refs}.  Despite
recent work by \citet{Frazier+Powell:2010}, the theory of selection
problems is less well understood than that of bandit problems.
Accordingly, we present in \secrefs{optimal-section}{context-section}
a number of results concerning optimal policies for the general case
and for specific sampling distributions. In particular, we give finite
bounds on the number of samples collected by optimal policies for
Bernoulli and Gaussian distributions. We also provide a simple
counterexample to the intuitive conjecture that an optimal policy
should not spend more on deciding than the decision is worth; in fact,
it is possible for an optimal policy to compute forever. We also show
by a simple counterexample that {\em index
policies}~\cite{Gittins:1989} do not necessarily exist for selection
problems.

>>>>

The broad family of sequential decision problems includes
combinatorial search problems, game playing, robotic path planning,
model-predictive control problems, Markov decision processes (MDP), whether fully or
partially observable, and a huge range of applications. In almost all
realistic instances, exact solution is intractable and approximate
methods are sought. Perhaps the most popular approach is to simulate
a limited number of possible future action sequences, 
in order to find a move in the current state that is (hopefully)
near-optimal. One of the most successful
examples of this approach are sampling algorithms for
playing the game of Go \citep{Gelly.mogo} based on the upper-confidence bound (UCB)
method \citep{Kocsis+Szepesvari:2006}. This paper examines the issue of sampling for move selection from
first principles, by defining an appropriate meta-level MDP and formally
analyzing some of its properties. Although
the meta-level MDP is also intractable, approximation schemes based on
provable value of information bounds for this problem are presented. 
These lead to practical sampling schemes that
are promising: they out-perform the UCB-based schemes,
as indicated by empirical results on both ``flat'' selection and
game-playing in Go.

As selection is extremely general \citep{TolpinShimony:2012}, we focus on the adversarial
game-playing domain as an example for clarity.
A typical game-playing algorithm explores a tree or graph of action sequences
with some limit on depth, using pruning methods to avoid irrelevant
subtrees; based on what it finds in the explored portion of the state space, the algorithm
then selects a move. This exploration is repeated after observing the move
made by the opponent (whether or not information from previous explorations
is reused is discussed in Section \ref{sec:control-redistribution}).

Clearly, it is desirable to select the best possible move using the
least possible amount of exploration. For a given amount of
exploration, decision quality can be improved by directing exploration
towards those actions sequences whose outcomes are helpful in selecting
a good move. Thus, the {\em metalevel} decision problem is to choose
what future action sequences to explore (or, more generally, what
deliberative computations to do), while the {\em object-level}
decision problem is to choose an action to execute in the real world.

@@

An independent thread of research on metalevel control began with work
by \citet{Kocsis+Szepesvari:2006} on the UCT algorithm, which operates
in the context of {\em Monte Carlo tree search} (MCTS) algorithms.
In MCTS, each computation takes the form
of a simulating a randomized sequence of actions leading from a leaf of the
current tree to a terminal state. UCT is primarily a method for
selecting a leaf from which to conduct the next simulation, and
forms the core of the successful \textsc{MoGo} algorithm for Go 
playing \citep{Gelly+Silver:2011}.  The UCT algorithm is
based on the the theory of bandit problems \citep{Berry+Fristedt:1985} and the near-optimal
UCB1 bandit algorithm \citep{Auer+et+al:2002}. UCT applies
UCB1 recursively to select actions to perform within simulations.

@@

Analyzing the metalevel problem within an appropriate theoretical
framework ought to lead to more effective algorithms than those
obtained within the bandit framework.  For Monte Carlo computations,
in which samples are gathered to estimate the utilities of actions,
the metalevel decision problem is an instance of the {\em selection
problem} studied in statistics~\citep{Bechhofer:1954,Swisher+et+al:2003}.  Despite
recent work by \citet{Frazier+Powell:2010} and \citet{TolpinShimony:2012}, the theory of selection
problems is less well understood than that of bandit problems.
Accordingly, we present in \secrefs{sec:optimal}{sec:context}
a number of results concerning optimal policies for the general case
and for specific sampling distributions. In particular, we give finite
bounds on the number of samples collected by optimal policies for
Bernoulli and Gaussian distributions. We also provide a simple
counterexample to the intuitive conjecture that an optimal policy
should not spend more on deciding than the decision is worth; in fact,
it is possible for an optimal policy to compute forever. We also show
by a simple counterexample that optimal {\em index
policies}~\citep{Gittins:1989} do not necessarily exist for selection
problems.
